---
title: "Pipeline Usage Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pipeline Usage Guide}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This guide shows how to process text responses through the SADCAT pipeline,
either as a single call or stage-by-stage for full control over each step.

## Dependencies at a glance

The core pipeline (preprocess, valence, dictionaries, aggregate) runs on a
standard R installation with no Java and no Python. Two parts are optional and
carry extra requirements:

- **Spell-checking in preprocessing.** Only relevant when `spellcheck = TRUE`.
  The default `"hunspell"` method needs Java and a WordNet dictionary; the
  `"gemini"` method needs a Gemini API key. Set `spellcheck = FALSE` to skip the
  step and require neither.
- **Embeddings and seed similarities (Stages 4 and 5).** Need Python through
  `reticulate` (SBERT uses the `sentence-transformers` library; Gemini
  embeddings use Google's `genai` library plus an API key). These stages are
  optional, and the dictionary-based coding is complete without them.

## Setup

```{r}
library(SADCAT)
```

## Input data format

Your data should be a data.frame with at least one text column containing the
raw responses. Additional columns (e.g., grouping variables) are preserved
through the pipeline.

```{r}
dat <- data.frame(
  responsex = c("warm and friendly", "cold and rde", "smart worker", NA),
  Target = c("A", "A", "B", "B"),
  Participant = c("1", "2", "3", "4"),
  stringsAsFactors = FALSE
)
```


Coding stages: `"preprocess"`, `"valence"`, `"dictionaries"`, `"socats"`,
`"embeddings"`, `"seeds"`, `"aggregate"`.


### Stage 1: Preprocess text

Cleans and normalizes text responses. Produces three text columns:

- **`tv`**: lowercased text (used for valence scoring)
- **`tv2`**: spell-checked text
- **`tv3`**: singularized text (used for dictionary matching)

Spell-checking is optional and is governed by two arguments: `spellcheck`
(whether to run it at all) and `spellcheck_method` (how to run it, when it
runs).

Set `spellcheck = FALSE` to skip spell-checking entirely. This requires neither
Java nor an API key, and is the simplest option if your responses are already
clean or you just want a first pass. When skipped, `tv2` is copied directly
from `tv`.

When `spellcheck = TRUE`, two methods are available via `spellcheck_method`:

- **`"hunspell"`** (default): Uses Java + WordNet + hunspell. Suggestions are
  ranked by edit distance, with bigram context as a tiebreaker when the corpus
  has 50+ unique values. Auto-detects Java and WordNet paths across
  macOS/Linux/Windows. If auto-detection fails, set explicit `java_home` /
  `wordnet_dict` arguments or environment variables (`JAVA_HOME`,
  `SADCAT_WORDNET_DICT`).
- **`"gemini"`**: **RECOMMENDED SPELL-CHECKING METHOD**.
  Uses the Gemini LLM for context-aware correction. Sends full
  responses (not individual words) so the model can use surrounding context.
  Requires a Gemini API key via `gemini_spellcheck_key` or the
  `GEMINI_API_KEY` environment variable. May require updating the model, check:
  https://ai.google.dev/gemini-api/docs/models. You can also provide a custom 
  context prompt via `gemini_spellcheck_context` to give the model more 
  information about the domain of your text data. The default prompt describes 
  the text as short responses describing a social target. 

```{r}
dat <- preprocess_text(
  dat,
  text_col = "responsex",
  spellcheck = TRUE,                     # FALSE skips it entirely (no Java or API key needed)
  spellcheck_method = "gemini",         # "hunspell" (default) or "gemini"
  singularize = TRUE,
  java_home = NULL,                       # optional; used for hunspell only
  wordnet_dict = NULL,                    # optional; used for hunspell only
  gemini_spellcheck_key = NULL,           # optional; used for gemini only
  gemini_spellcheck_model = "gemini-3.1-flash-lite-preview",
  gemini_spellcheck_context = NULL,       # optional custom Gemini prompt
  verbose = TRUE
)

# Inspect: what did preprocessing change?
dat[, c("responsex", "tv", "tv2", "tv3")]
```



### Stage 2: Score valence

Scores each response against 5 sentiment dictionaries (Lexicoder/LSD2015,
NRC, Bing, AFINN, Loughran). Produces per-dictionary scores and combined
scores with negation handling.

**Key output columns:**


- **`ValenceYesNA`**: Combined valence. Mean of negation-aware scores across
  the 5 dictionaries. NA when no dictionary matched any sentiment words.
  Negation is applied once to this combined score.
- **`ValenceNoNA`**: Same as `ValenceYesNA`, but 0 instead of NA.

```{r}
dat <- score_valence(
  dat,
  text_col = "tv2",             # recommend using "tv2" for human participants
  response_col = "responsex"    # NA-gating column; NULL defaults to text_col
)

# Inspect valence
dat[, c("responsex", "Val_lexicoder", "Val_NRC", "Val_bing", "Val_affin", "Val_loughran", "ValenceYesNA", "ValenceNoNA")]
```

### Stage 3: Match SADCAT dictionaries

Matches text against the SADCAT stereotype content dictionaries and computes
binary indicators, percentages, direction scores, and per-dimension valence.

**Key output columns (per dimension, e.g., Warmth, Competence, Morality, ...):**

- `{Dim}_prevalence`: 1 if any match, 0 if none, NA if response missing
- `{Dim}_Valence` *(default for downstream means)*: NA if dimension absent;
  otherwise the global `ValenceNoNA` (0 for sentiment-less tagged responses,
  signed value otherwise). Use `mean(., na.rm = TRUE)` to average over tagged
  responses where sentiment-less ones contribute 0 (neutral).
- `{Dim}_valenceStrictNA`: NA if dimension absent **or** `ValenceYesNA` is NA.
  Strictly NA-gated on both axes.
- `{Dim}_valenceNoNA`: 0 if dimension absent or no sentiment matched;
  NA only if response is missing.
- `{Dim}_direction` (directional dimensions only): hi-lo direction, NA when not applicable
- `{Dim}_directionNoNA` (directional dimensions only): same as `{Dim}_direction`, but 0 if dimension absent (still NA if response missing)
- `NoMatch`: 1 if no SADCAT dimension matched at all (NA if response missing)
-  Note: Warmth = Sociability + Morality; Competence = Ability + Assertiveness. Beauty is a subdimension of Appearance. Words may fall into multiple dictionaries.

```{r}
dat <- match_dictionaries(
  dat,
  text_col = "tv3",
  response_col = "responsex",   # NA-gating column; NULL defaults to text_col
  valence_col = "ValenceYesNA",
  valence_nona_col = "ValenceNoNA",
  sadcat_dict = NULL,           # NULL auto-loads SADCAT dictionaries
  socats_dict = NULL,           # NULL auto-loads SOCATS only if socats = TRUE
  socats = TRUE                 # include Social categories dictionaries as well (optional; default FALSE)
)

# Inspect dimension-level results
View(dat)

#Inspect top 20 words matching each dimension (for sanity check and to understand what the dictionaries are capturing)
get_top_dictionary_matches(dat, text_col = "tv3", top_n = 20)
```


### Stage 4 (optional): Compute embeddings

Generates sentence embeddings via SBERT and/or Gemini. Requires Python
via reticulate.

```{r}
dat <- compute_embeddings(
  dat,
  text_col = "tv",
  response_col = "responsex",   # NA-gating column; NULL defaults to text_col
  methods = c("sbert", "gemini"),           # or "sbert" if gemini not set up
  sbert_model = "paraphrase-mpnet-base-v2",
  sbert_dims = 768L,
  gemini_api_key = NULL,       # required only when methods includes "gemini" and key not set via env var
  gemini_model = "gemini-embedding-001",
  gemini_dims = 768L,
  gemini_batch_size = 10L,
  gemini_sleep = 63,
  gemini_task_type = "SEMANTIC_SIMILARITY",
  verbose = TRUE
)
```

### Stage 5 (optional): Compute seed similarities

Computes correlation (or cosine similarity) between response embeddings and
SADCAT seed vectors. Requires embeddings from Stage 4.

```{r}
dat <- compute_seed_similarities(
  dat,
  embedding_prefix = c("SBERT", "Gemini"),  # or only one model, based on stage 4 decisions
  seed_vectors = Seed_Vectors_Avg,
  method = "correlation",      # or "cosine"
  response_col = "responsex",  # NA-gating column
  verbose = TRUE
)
```

### Stage 6: Aggregate responses

Aggregates response-level data to group-level, using sum, mean, and distinct
operations. Auto-detects standard columns by naming patterns.

```{r}
agg <- aggregate_responses(
  dat,
  group_cols = c("Participant", "Target"),
  sum_cols = NULL,             # NULL = auto-detect
  mean_cols = NULL,            # NULL = auto-detect
  distinct_cols = NULL,        # NULL = auto-detect
  extra_sum_cols = NULL,
  extra_mean_cols = NULL,
  extra_distinct_cols = NULL,
  verbose = TRUE
)

View(agg)
```


