---
title: "Pipeline Usage Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pipeline Usage Guide}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This guide shows how to process text responses through the SADCAT pipeline,
either as a single call or stage-by-stage for full control over each step.

## Setup

```{r}
library(SADCAT)
```

## Input data format

Your data should be a data.frame with at least one text column containing the
raw responses. Additional columns (e.g., grouping variables) are preserved
through the pipeline.

```{r}
dat <- data.frame(
  responsex = c("warm and friendly", "cold and rde", "smart worker"),
  Synonym.GroupX = c("A", "A", "B"),
  Group = c("G1", "G1", "G1"),
  Level = c("L1", "L1", "L1"),
  stringsAsFactors = FALSE
)
```


Coding stages: `"preprocess"`, `"valence"`, `"dictionaries"`, `"socats"`,
`"embeddings"`, `"seeds"`, `"aggregate"`.


### Stage 1: Preprocess text

Cleans and normalizes text responses. Produces three text columns:

- **`tv`**: lowercased text (used for valence scoring)
- **`tv2`**: spell-checked text
- **`tv3`**: singularized text (used for dictionary matching)

Two spell-checking methods are available via the `spellcheck_method` parameter:

- **`"hunspell"`** (default): Uses Java + WordNet + hunspell. Suggestions are
  ranked by edit distance, with bigram context as a tiebreaker when the corpus
  has 50+ unique values. Auto-detects Java and WordNet paths across
  macOS/Linux/Windows. If auto-detection fails, set explicit `java_home` /
  `wordnet_dict` arguments or environment variables (`JAVA_HOME`,
  `SADCAT_WORDNET_DICT`).
- **`"gemini"`**: **RECOMMENDED SPELL-CHECKING METHOD**.
  Uses the Gemini LLM for context-aware correction. Sends full
  responses (not individual words) so the model can use surrounding context.
  Requires a Gemini API key via `gemini_spellcheck_key` or the
  `GEMINI_API_KEY` environment variable. May require updating the model, check:
  https://ai.google.dev/gemini-api/docs/models. You can also provide a custom 
  context prompt via `gemini_spellcheck_context` to give the model more 
  information about the domain of your text data. The default prompt describes 
  the text as short responses describing a social target. 

```{r}
dat <- preprocess_text(
  dat,
  text_col = "responsex",
  spellcheck = TRUE,                     # TRUE enables spell-checking
  spellcheck_method = "gemini",         # "hunspell" (default) or "gemini"
  singularize = TRUE,
  java_home = NULL,                       # optional; used for hunspell only
  wordnet_dict = NULL,                    # optional; used for hunspell only
  gemini_spellcheck_key = NULL,           # optional; used for gemini only
  gemini_spellcheck_model = "gemini-3.1-flash-lite-preview",
  gemini_spellcheck_context = NULL,       # optional custom Gemini prompt
  verbose = TRUE
)

# Inspect: what did preprocessing change?
dat[, c("responsex", "tv", "tv2", "tv3")]
```



### Stage 2: Score valence

Scores each response against 5 sentiment dictionaries (Lexicoder/LSD2015,
NRC, Bing, AFINN, Loughran). Produces per-dictionary scores and combined
scores with negation handling.

**Key output columns:**


- **`Valy`**: Combined valence. Mean of negation-aware scores across the
  5 dictionaries. NA when no dictionary matched any sentiment words.
  Negation is applied once to this combined score.
- **`ValyNoNA`**: Same as `Valy`, but 0 instead of NA.

```{r}
dat <- score_valence(
  dat,
  text_col = "tv2",             # recommend using "tv2" for human participants
  response_col = "responsex"    # NA-gating column; NULL defaults to text_col
)

# Inspect valence
dat[, c("responsex", "Val_lexicoder", "Val_NRC", "Val_bing", "Val_affin", "Val_loughran", "Valy", "ValyNoNA")]
```

### Stage 3: Match SADCAT dictionaries

Matches text against the SADCAT stereotype content dictionaries and computes
binary indicators, percentages, direction scores, and per-dimension valence.

**Key output columns (per dimension, e.g., Warmth, Competence, Morality, ...):**

- `_dic`: raw word count matching the dimension
- `_dic_binary`: 1 if any match, 0 otherwise
- `_dic_binary2`: same, but NA if the response is missing **use this for prevalence**
- `_dirx`: direction score (hi - lo binary), negation-aware
- `_Valy`: global `Valy` gated by dimension (NA if dimension not present)
- `_ValyNoNA`: global `ValyNoNA` gated by dimension (0 if dimension not present)
- `_valy3`: `_Valy` further gated by response presence **use this for valence**
- `_dirx3`: `_dirx` gated by binary and binary2 (NA when not applicable) **use this for direction**
- `_dirx3NoNA`: `_dirx3` with 0 when dimension absent (NA only if response missing) **use this for direction with neutral coding of non-present dimensions**
- `_valyNoNA3`: `_ValyNoNA` gated by response presence only 
  (0 if dimension absent; NA if response missing) **use this for valence if many dimension NAs and willing to accept assumption of neutral valence**
- `None2y`: 1 if no dimension matched at all

```{r}
dat <- match_dictionaries(
  dat,
  text_col = "tv3",
  response_col = "responsex",   # NA-gating column; NULL defaults to text_col
  valence_col = "Valy",
  valence_nona_col = "ValyNoNA",
  sadcat_dict = NULL,           # NULL auto-loads SADCAT dictionaries
  socats_dict = NULL,           # NULL auto-loads SOCATS only if socats = TRUE
  socats = FALSE
)

# Inspect dimension-level results
dat[, c("responsex", "warmth_dic_binary", "Warmth_dirx", "Warmth_Valy", "Warmth_ValyNoNA")]
```

### Stage 3b (optional): Match SOCATS dictionaries

Matches against social category dictionaries (age, gender, race, etc.).

```{r}
dat <- match_dictionaries(
  dat,
  text_col = "tv3",
  response_col = "responsex",   # NA-gating column; NULL defaults to text_col
  valence_col = "Valy",
  valence_nona_col = "ValyNoNA",
  sadcat_dict = FALSE,          # skip SADCAT (already done)
  socats_dict = NULL,           # NULL auto-loads SOCATS dictionaries
  socats = TRUE
)
```

### Stage 4 (optional): Compute embeddings

Generates sentence embeddings via SBERT and/or Gemini. Requires Python
via reticulate.

```{r}
dat <- compute_embeddings(
  dat,
  text_col = "tv",
  response_col = "responsex",   # NA-gating column; NULL defaults to text_col
  methods = "sbert",           # or c("sbert", "gemini")
  sbert_model = "paraphrase-mpnet-base-v2",
  sbert_dims = 768L,
  gemini_api_key = NULL,       # required only when methods includes "gemini"
  gemini_model = "gemini-embedding-exp-03-07",
  gemini_dims = 2000L,
  gemini_batch_size = 10L,
  gemini_sleep = 63,
  gemini_task_type = "SEMANTIC_SIMILARITY",
  verbose = TRUE
)
```

### Stage 5 (optional): Compute seed similarities

Computes correlation (or cosine similarity) between response embeddings and
SADCAT seed vectors. Requires embeddings from Stage 4.

```{r}
dat <- compute_seed_similarities(
  dat,
  embedding_prefix = "SBERT",  # or "Gemini"
  seed_vectors = Seed_Vectors_Avg,
  method = "correlation",      # or "cosine"
  response_col = "responsex",  # NA-gating column
  verbose = TRUE
)
```

### Stage 6: Aggregate responses

Aggregates response-level data to group-level, using sum, mean, and distinct
operations. Auto-detects standard columns by naming patterns.

```{r}
agg <- aggregate_responses(
  dat,
  group_cols = c("Synonym.GroupX", "Group", "Level"),
  sum_cols = NULL,             # NULL = auto-detect
  mean_cols = NULL,            # NULL = auto-detect
  distinct_cols = NULL,        # NULL = auto-detect
  extra_sum_cols = NULL,
  extra_mean_cols = NULL,
  extra_distinct_cols = NULL,
  create_noNA = TRUE,
  verbose = TRUE
)

head(agg)
```

## Valence column summary

| Column | Level | NA behavior | Negation |
|--------|-------|-------------|----------|
| `Val_lexicoder`, `Val_NRC`, ... | Per-dictionary | 0 when no match | Raw (no flip) |
| `Val_lexicoderNA`, `Val_NRCNA`, ... | Per-dictionary | NA when no match | Raw (no flip) |
| `Valy` | Combined (5-dict mean) | NA when no dict matched | Applied once |
| `ValyNoNA` | Combined (5-dict mean) | 0 when no dict matched | Applied once |
| `{Dim}_Valy` | Per-dimension | NA when dimension not present | Inherited |
| `{Dim}_ValyNoNA` | Per-dimension | 0 when dimension not present | Inherited |
| `{Dim}_valy3` | Per-dimension, gated | NA when dim or response missing | Inherited |
| `{Dim}_valyNoNA3` | Per-dimension, gated | NA when response missing; 0 when dim missing | Inherited |

## Variables reference

- **`_dic`**: raw word count for the dimension
- **`_dic_hi` / `_dic_lo`**: word counts for high- and low-directional words
- **`_dic_percent`**: proportion of tokens matching the dimension
- **`_dic_binary`**: 1 if any words match, 0 otherwise
- **`_dic_binary2`**: same as `_binary`, but NA if original response is missing
- **`_dirx`**: direction score: `hi_binary - lo_binary` (negation-aware)
- **`_dirx2`**: same as `_dirx` (negation-aware version)
- **`_dirx3`**: `_dirx` gated by binary and binary2 (NA when not applicable)
- **`_dirx3NoNA`**: `_dirx3` with 0 when dimension is absent (NA only if response is missing)
- **`None2y`**: 1 if no dimension matched at all, NA if response missing
