---
title: "Pipeline Usage Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pipeline Usage Guide}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This guide shows how to process text responses through the SADCAT pipeline,
either as a single call or stage-by-stage for full control over each step.

## Setup

```{r}
library(SADCAT)
```

## Input data format

Your data should be a data.frame with at least one text column containing the
raw responses. Additional columns (e.g., grouping variables) are preserved
through the pipeline.

```{r}
dat <- data.frame(
  responsex = c("warm and friendly", "cold and rde", "smart worker"),
  Synonym.GroupX = c("A", "A", "B"),
  Group = c("G1", "G1", "G1"),
  Level = c("L1", "L1", "L1"),
  stringsAsFactors = FALSE
)
```

## Option A: All-in-one with `process_responses()`

The simplest way to run the full pipeline. Returns a list with `$long`
(response-level) and `$agg` (group-level aggregated) data.

Two spell-checking methods are available via the `spellcheck_method` parameter:

- **`"hunspell"`** (default): Uses Java + WordNet + hunspell. Suggestions are
  ranked by edit distance, with bigram context as a tiebreaker when the corpus
  has 50+ unique values. Auto-detects Java and WordNet paths across
  macOS/Linux/Windows. If auto-detection fails, set explicit `java_home` /
  `wordnet_dict` arguments or environment variables (`JAVA_HOME`,
  `SADCAT_WORDNET_DICT`).
- **`"gemini"`**: Uses the Gemini LLM for context-aware correction. Sends full
  responses (not individual words) so the model can use surrounding context.
  Requires a Gemini API key via `gemini_spellcheck_key` or the
  `GEMINI_API_KEY` environment variable.

```{r}
result <- process_responses(
  dat,
  text_col = "responsex",
  group_cols = c("Synonym.GroupX", "Group", "Level"),
  stages = c("preprocess", "valence", "dictionaries", "aggregate"),
  spellcheck = FALSE,          # set TRUE for spell-checking
  singularize_text = TRUE,
  verbose = TRUE
)

head(result$long)
head(result$agg)
```

Use the `stages` parameter to selectively run only the stages you need.
Available stages: `"preprocess"`, `"valence"`, `"dictionaries"`, `"socats"`,
`"embeddings"`, `"seeds"`, `"aggregate"`.

## Option B: Stage-by-stage

Running each stage separately lets you inspect intermediate results, customize
parameters per stage, or integrate with your own processing steps.

### Stage 1: Preprocess text

Cleans and normalizes text responses. Produces three text columns:

- **`tv`**: lowercased text (used for valence scoring)
- **`tv2`**: spell-checked text
- **`tv3`**: singularized text (used for dictionary matching)

```{r}
dat <- preprocess_text(
  dat,
  text_col = "responsex",
  spellcheck = FALSE,          # TRUE enables spell-checking (see below)
  singularize = TRUE,
  verbose = TRUE
)

# Inspect: what did preprocessing change?
dat[, c("responsex", "tv", "tv2", "tv3")]
```

#### Hunspell spell-checking (default)

Uses Java + WordNet + hunspell. Suggestions are ranked by edit distance, with
bigram context from the corpus used as a tiebreaker (auto-triggered when 50+
unique values are present). If auto-detection of Java/WordNet fails, override
explicitly:

```{r}
dat <- preprocess_text(
  dat,
  text_col = "responsex",
  spellcheck = TRUE,
  spellcheck_method = "hunspell",    # default
  java_home = "/path/to/java/home",
  wordnet_dict = "/path/to/wordnet/dict",
  verbose = TRUE
)
```

#### Gemini LLM spell-checking (optional)

Sends the full response texts to Gemini for context-aware correction. This
avoids the need for Java/WordNet and can produce better results by using
surrounding words as context. Requires a Gemini API key.

```{r}
dat <- preprocess_text(
  dat,
  text_col = "responsex",
  spellcheck = TRUE,
  spellcheck_method = "gemini",
  gemini_spellcheck_key = "your-api-key",    # or set GEMINI_API_KEY env var
  gemini_spellcheck_model = "gemini-2.0-flash",
  verbose = TRUE
)
```

You can also provide a custom context prompt via `gemini_spellcheck_context`
to give the model more information about the domain of your text data.
The default prompt describes the text as short responses describing a social
target.

### Stage 2: Score valence

Scores each response against 5 sentiment dictionaries (Lexicoder/LSD2015,
NRC, Bing, AFINN, Loughran). Produces per-dictionary scores and combined
scores with negation handling.

**Key output columns:**

- `Val_lexicoder`, `Val_NRC`, `Val_bing`, `Val_affin`, `Val_loughran`:
  Per-dictionary raw scores (-1, 0, or 1). These are raw presence indicators
  with no negation flipping.
- `Val_lexicoderNA`, `Val_NRCNA`, etc.: Same, but NA when that dictionary
  found no sentiment words.
- **`Valy`**: Combined valence. Mean of negation-aware scores across the
  5 dictionaries. NA when no dictionary matched any sentiment words.
  Negation is applied once to this combined score (not to individual
  dictionary scores).
- **`ValyNoNA`**: Same as `Valy`, but 0 instead of NA.

```{r}
dat <- score_valence(
  dat,
  text_col = "tv"              # use "tv2" for human participants
)

# Inspect valence
dat[, c("responsex", "Val_lexicoder", "Val_NRC", "Val_bing", "Valy", "ValyNoNA")]
```

### Stage 3: Match SADCAT dictionaries

Matches text against the SADCAT stereotype content dictionaries and computes
binary indicators, percentages, direction scores, and per-dimension valence.

**Key output columns (per dimension, e.g., Warmth, Competence, Morality, ...):**

- `warmth_dic`: raw word count matching the dimension
- `warmth_dic_binary`: 1 if any match, 0 otherwise
- `warmth_dic_binary2`: same, but NA if the response is missing
- `Warmth_dirx`: direction score (hi - lo binary), negation-aware
- `Warmth_Valy`: global `Valy` gated by dimension (NA if dimension not present)
- `Warmth_ValyNoNA`: global `ValyNoNA` gated by dimension (0 if dimension not present)
- `Warmth_valy3`: `Warmth_Valy` further gated by response presence
- `Warmth_valyNoNA3`: `Warmth_ValyNoNA` further gated by response presence
- `None2y`: 1 if no dimension matched at all

```{r}
dat <- match_dictionaries(
  dat,
  text_col = "tv3",
  valence_col = "Valy",
  valence_nona_col = "ValyNoNA"
)

# Inspect dimension-level results
dat[, c("responsex", "warmth_dic_binary", "Warmth_dirx", "Warmth_Valy", "Warmth_ValyNoNA")]
```

### Stage 3b (optional): Match SOCATS dictionaries

Matches against social category dictionaries (age, gender, race, etc.).

```{r}
dat <- match_dictionaries(
  dat,
  text_col = "tv3",
  valence_col = "Valy",
  valence_nona_col = "ValyNoNA",
  sadcat_dict = FALSE,          # skip SADCAT (already done)
  socats = TRUE
)
```

### Stage 4 (optional): Compute embeddings

Generates sentence embeddings via SBERT and/or Gemini. Requires Python
via reticulate.

```{r}
dat <- compute_embeddings(
  dat,
  text_col = "tv",
  methods = "sbert",           # or c("sbert", "gemini")
  sbert_model = "paraphrase-mpnet-base-v2"
)
```

### Stage 5 (optional): Compute seed similarities

Computes correlation (or cosine similarity) between response embeddings and
SADCAT seed vectors. Requires embeddings from Stage 4.

```{r}
dat <- compute_seed_similarities(
  dat,
  embedding_prefix = "SBERT",  # or "Gemini"
  method = "correlation"
)
```

### Stage 6: Aggregate responses

Aggregates response-level data to group-level, using sum, mean, and distinct
operations. Auto-detects standard columns by naming patterns.

```{r}
agg <- aggregate_responses(
  dat,
  group_cols = c("Synonym.GroupX", "Group", "Level"),
  create_noNA = TRUE,
  verbose = TRUE
)

head(agg)
```

## Valence column summary

| Column | Level | NA behavior | Negation |
|--------|-------|-------------|----------|
| `Val_lexicoder`, `Val_NRC`, ... | Per-dictionary | 0 when no match | Raw (no flip) |
| `Val_lexicoderNA`, `Val_NRCNA`, ... | Per-dictionary | NA when no match | Raw (no flip) |
| `Valy` | Combined (5-dict mean) | NA when no dict matched | Applied once |
| `ValyNoNA` | Combined (5-dict mean) | 0 when no dict matched | Applied once |
| `{Dim}_Valy` | Per-dimension | NA when dimension not present | Inherited |
| `{Dim}_ValyNoNA` | Per-dimension | 0 when dimension not present | Inherited |
| `{Dim}_valy3` | Per-dimension, gated | NA when dim or response missing | Inherited |
| `{Dim}_valyNoNA3` | Per-dimension, gated | NA when dim or response missing | Inherited |

## Variables reference

- **`_dic`**: raw word count for the dimension
- **`_dic_hi` / `_dic_lo`**: word counts for high- and low-directional words
- **`_dic_percent`**: proportion of tokens matching the dimension
- **`_dic_binary`**: 1 if any words match, 0 otherwise
- **`_dic_binary2`**: same as `_binary`, but NA if original response is missing
- **`_dirx`**: direction score: `hi_binary - lo_binary` (negation-aware)
- **`_dirx2`**: same as `_dirx` (negation-aware version)
- **`_dirx3`**: `_dirx` gated by binary and binary2 (NA when not applicable)
- **`None2y`**: 1 if no dimension matched at all, NA if response missing
